Thesis: draft

### 2021-05-18 draft.tex
added results for PGP trust network

Het is gekend dat mensen niet enkel hun mening uit zichzelf vormen, maar sterk beïnvloedt worden door onder andere de mensen rond hen, hun omgeving, hun opvoeding,... Sociale media laten onmiddellijke interacties toe tussen mensen van overal op de wereld, zonder daarbij belemmerd te worden door fysieke en/of tijdsgrenzen. Dit kan een potentieel grote impact hebben op hoe meningen zich vormen en hoe ze evolueren.\\

@@

\selectlanguage{english}
\tableofcontents
\addcontentsline{toc}{chapter}{Table of Contents}
\newgeometry{top=20mm, bottom=25mm}
\listoffigures
\addcontentsline{toc}{chapter}{List of Figures}
\listoftables
\addcontentsline{toc}{chapter}{List of Tables}
\restoregeometry

@@

The results of the two real-world networks: (i) the Last.fm on-line social network; and (ii) the PGP trust network, discussed in Chapter \ref{chap4}, Section \ref{realnetwork}, will be shared and discussed. For each real-world network, the results will be compared to results obtained from network models with the same average degree, clustering and/or modularity. We will mainly focus on the formation of echo chambers. We will only study balanced initial conditions and we will introduce a fraction of completely stubborn actors. For both real-world networks, the REC and PR filtering algorithms will be compared.

@@

The formation of echo chambers versus the fraction of completely stubborn actors, fracRes, in the Last.fm on-line social network is studied. The results are compared to corresponding results from network models, all with $N = 8003$: (i) the Watts-Strogatz model with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) the stochastic block model with 53 communities of size 151, $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$; (iii) the stochastic Watts-Strogatz block model with 53 communities of size 151, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$. The results for the PR and REC filtering algorithms are shown in Fig. \ref{echo_vs_fracRes_real}.\\
\newline
The PR filtering algorithm shows a decrease in the formation of echo chambers with an increasing fraction of stubborn actors. This is valid for both the real-world network and the network models. The stochastic Watts-Strogatz block model has similar clustering and modularity as the Last.fm network. The Last.fm network has, however, a smaller formation of echo chambers than the SBM-WS. This suggests that it is not only clustering and/or modularity that drives the formation of echo chambers. Other topologies must play a role as well. The biggest difference between the real-world Last.fm network and the stochastic Watts-Strogatz block model lies in the degree distribution: the Last.fm network has a heterogeneous degree distribution instead of a homogeneous degree distribution. It is probably this heterogeneous contact pattern that hampers the formation of echo chambers in the real-world network compared to a network model with similar clustering and modularity, but with a homogeneous degree distribution. This agrees with the result that Perra and Rocha \cite{Perra2019} found: a heterogeneous contact pattern hampers the formation of echo chambers. Another difference between the Last.fm and the SBM and SBM-WS network is the community structure. The Last.fm network has a community structure where the community sizes follow a heterogeneous distribution: plenty of small communities and a few large ones. The stochastic block model and stochastic Watts-Strogatz block model, that we used, only have communities of a single size. This does, however, not play a significant role in the formation of echo chambers, as can be seen in Fig. \ref{power_vs_single}. In this figure, the distribution of echo chambers in the Last.fm network is compared to those in the SBM-WS network with similar clustering and modularity. The SBM-WS is modeled in two ways: (i) the communities all have a fixed size; (ii) the communities have the same heterogeneous sizes as in the Last.fm network. The PR and REC filtering algorithms are compared.\\ % add parameters of SBM-WS power??
\newpage
\noindent
The results for the Watts-Strogatz model are similar to those for the stochastic Watts-Strogatz block model. The formation of echo chambers seems to be slightly smaller in the WS than in the SBM-WS network. It may be tempting to attribute this to the lack of community structure in the WS model. It must however be noted that the implemented WS model has a slightly smaller clustering coefficient ($\left<cc\right> = 0.317 \pm 0.004$) than the SBM-WS ($\left<cc\right> = 0.344 \pm 0.002$), which may also explain the somewhat smaller formation of echo chambers. Furthermore, it is interesting to notice that the Last.fm on-line social network coincides with the SBM with a similar modularity. It seems that the hampering of the formation of echo chambers due to a heterogeneous contact pattern is of the same size as the hampering due to a lack of high clustering. % talk about coincidence??

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{PR}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_PR_real_network_8x7.png}
    \label{pr_real}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{REC}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_REC_real_network_8x7.png}
    \label{rec_real}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $50/50$ opinion distribution. The Last.fm on-line social network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $50/50$ opinion distribution. The Last.fm on-line social network is compared to corresponding network models: (i) WS with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) SBM-WS with 53 communities of size 151, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$; (iii) SBM with 53 communities of size 151 with $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$. \textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Each data point is the average of $50$ independent simulations. Bars represent the standard deviation $\sigma$.}
\label{echo_vs_fracRes_real}
\end{figure}
\noindent
The REC filtering algorithm enforces an increase in the formation of echo chambers when a small fraction of stubborn actors is introduced. If this fraction is further increased, the formation of echo chambers starts to decrease. The increase is the strongest for the high clustering network models (WS, SBM-WS). Even though the Last.fm network has a similar clustering as these high clustering models, the formation of echo chambers and the corresponding increase in that formation are significantly smaller. As for the PR filtering algorithm, this is probably a consequence of the heterogeneous degree pattern of the Last.fm network. The results for the Last.fm network coincide with those of the SBM. As for the PR filtering algorithm, this seems to suggest that the hampering in the formation of echo chambers due to a heterogeneous degree distribution is of a similar size as the corresponding hampering due to the lack of high clustering.

@@

Results for the PGP trust network are compared to results for network models with similar average degree, clustering and/or modularity. The used network models are, all with $N = 10680$: (i) the Watts-Strogatz model with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) the stochastic block model with 60 communities of size 178, $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$; (iii) the stochastic Watts-Strogatz block model with 60 communities of size 178, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$.

@@

It must be emphasized that there are limitations to our model and methods. As already stated before, the simpleness of the model, which on one hand can be seen as a strength, may also be one of the weaknesses. Real-life social behavior and opinion dynamics on social media may be far from simple. As a consequence, there may be many dynamics of real-life opinion formation that are not well-captured by our model. More empirical data is therefore needed to compare the results, obtained from our model, against. Other limitations may be the rather small system sizes that were used in our simulations ($N = 10^3$). It is unclear how the system size affects the results and it is recommended that the model is tested on both larger network models and real-world networks. Another possible shortcoming is the binary nature of the model. There are real-life cases in which people have to choose between two competing opinions. These situations, however, do not cover the complete range of possible opinion dynamics situations. It would be interesting to run the model on situations where more than two opinions are allowed or to switch to continuous opinion dynamics. We also did not differentiate between the agents in our model, except for non-stubborns versus stubborns. This combined with the homogeneous degree distribution of our networks models may be another limitation. Real-life social media platforms display a heterogeneous contact pattern, where the `hubs' are often more influential than other users, not only by their larger reach, but also by their status. This may impact the spread of opinions \cite{Fan2020}\cite{Wang2019}. We could adapt our model to mimic this real social media behavior more closely. Therefore, a heterogeneous network model should be introduced, combined with an influence parameter. This parameter should be used to determine the influence of an agent on others: the opinion of a more influential agent is easier adopted by the others. In the model, the high degree nodes should then be made more influential. As a final limitation, we mention the fixed stubbornness of the agents. Even though the stubbornness parameter $r$ could be individualized to each node, we did not include the possibility to be more stubborn towards opinions coming from one friend than from another. This idea may be linked to the influence of certain friends/acquaintances: somebody may adopt another viewpoint more easily from certain friends than from others.\\

@@

Simple toy models were implemented to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions. Our goal was to get a better understanding of the impact of social media and algorithmic personalization on phenomena such as polarization and formation of echo chambers (opinion clusters) in society. \\
\newline
We found that topological correlations such as high clustering strongly enhance the formation of opinion clusters. Beside clustering, community structure also favors the formation of echo chambers. Real-world networks have a lower formation of echo chambers than network models with similar clustering and modularity. The heterogeneous contact pattern of these real-world networks is mostly likely the cause of this hampering. If people are not stubborn, we found that semantic filtering and hence, a homogeneous exposure to information increases polarization in the system. This is an intuitive result: if you shield people from opposing views and strengthen them in their own beliefs, it feels natural that the system will evolve towards a more polarized state. It is then expected that a more heterogeneous exposure to information reduces this polarization. We found that this is indeed the case if people are not stubborn. If, on the contrary, people have some level of stubbornness, we found the surprising result that homogeneous exposure to information may be more protective against opinion clusters than heterogeneous exposure. This is an important result, since it may change our beliefs/assumptions that algorithmic personalization is at the root of the increase in polarization and extremism often seen on social media. More research on this topic is definitely needed in the near future. Finally, we found that distributing the opinions in communities instead of randomly distributing them may be hinder the convergence towards the majority opinion in case of semantic filtering, if the modularity is high enough.\\
\newline
Our results strongly suggest that the effects of algorithmic personalization do not stand on their own. Instead, the interplay with certain network topologies may enhance or hamper these effects. Beside that, individual resistance strongly affects the impact of these filtering algorithms. It would therefore be incorrect to only point to algorithmic personalization as the cause of polarization. It is the interplay between many different aspects of the system (such as topological correlations, distribution of the opinions, individual resistance, content curration) that may potentially enhance or hamper the formation of opinion clusters.
% also discuss need to investigate effect of changing N, S and p_act
\newpage

>>>>

Het is gekend dat mensen niet enkel hun mening uit zichzelf vormen, maar sterk beïnvloed worden door onder andere de mensen rond hen, hun omgeving, hun opvoeding,... Sociale media laten onmiddellijke interacties toe tussen mensen van overal op de wereld, zonder daarbij belemmerd te worden door fysieke en/of tijdsgrenzen. Dit kan een potentieel grote impact hebben op hoe meningen zich vormen en hoe ze evolueren.\\

@@

\selectlanguage{english}
\tableofcontents
\addcontentsline{toc}{chapter}{Table of Contents}
%\newgeometry{top=20mm, bottom=25mm}
\listoffigures
\addcontentsline{toc}{chapter}{List of Figures}
\listoftables
\addcontentsline{toc}{chapter}{List of Tables}
%\restoregeometry

@@

The results of the two real-world networks: (i) the Last.fm on-line social network; and (ii) the PGP trust network, discussed in Chapter \ref{chap4}, Section \ref{realnetwork}, will be shared and discussed. For each real-world network, the results will be compared to results obtained from network models with similar average degree, clustering and/or modularity. We will mainly focus on the formation of echo chambers. We will only study balanced initial conditions and we will introduce a fraction of completely stubborn actors. For both real-world networks, the REC and PR filtering algorithms will be compared.

@@

The formation of echo chambers versus the fraction of completely stubborn actors, fracRes, in the Last.fm on-line social network is studied. The results are compared to corresponding results from network models, all with $N = 8003$: (i) the Watts-Strogatz model with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) the stochastic block model with 53 communities of size 151, $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$; (iii) the stochastic Watts-Strogatz block model with 53 communities of size 151, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$. The results for the PR and REC filtering algorithms are shown in Fig. \ref{echo_vs_fracRes_real}.\\
\newline
The PR filtering algorithm shows a decrease in the formation of echo chambers with an increasing fraction of stubborn actors. This is valid for both the real-world network and the network models. The stochastic Watts-Strogatz block model has similar clustering and modularity as the Last.fm network. The Last.fm network has, however, a smaller formation of echo chambers than the SBM-WS. This suggests that it is not only clustering and/or modularity that drives the formation of echo chambers. Other topologies must play a role as well. The biggest difference between the real-world Last.fm network and the stochastic Watts-Strogatz block model lies in the degree distribution: the Last.fm network has a heterogeneous degree distribution instead of a homogeneous degree distribution. It is probably this heterogeneous contact pattern that hampers the formation of echo chambers in the real-world network compared to a network model with similar clustering and modularity, but with a homogeneous degree distribution. This agrees with the result that Perra and Rocha \cite{Perra2019} found: a heterogeneous contact pattern hampers the formation of echo chambers.\\
\newline 
Another difference between the Last.fm and the SBM and SBM-WS network is the community structure. The Last.fm network has a community structure where the community sizes follow a heterogeneous distribution: plenty of small communities and a few large ones. The stochastic block model and stochastic Watts-Strogatz block model, that we used, only have communities of a single size. This does, however, not play a significant role in the formation of echo chambers, as can be seen in Fig. \ref{power_vs_single}. In this figure, the distribution of echo chambers in the Last.fm network is compared to those in the SBM-WS network with similar clustering and modularity. The SBM-WS is modeled in two ways: (i) the communities all have a fixed size; (ii) the communities have the same heterogeneous sizes as in the Last.fm network. The PR and REC filtering algorithms are compared.\\ % add parameters of SBM-WS power??
\newpage
\noindent
The results for the Watts-Strogatz model are similar to those for the stochastic Watts-Strogatz block model. The formation of echo chambers seems to be slightly smaller in the WS than in the SBM-WS network. It may be tempting to attribute this to the lack of community structure in the WS model. It must however be noted that the implemented WS model has a slightly smaller clustering coefficient ($\left<cc\right> = 0.317 \pm 0.004$) than the SBM-WS ($\left<cc\right> = 0.344 \pm 0.002$), which may also explain the somewhat smaller formation of echo chambers. Furthermore, it is interesting to notice that the Last.fm on-line social network coincides with the SBM with a similar modularity. It seems that the hampering of the formation of echo chambers due to a heterogeneous contact pattern is of the same size as the hampering due to a lack of high clustering. % talk about coincidence??

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{PR}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_PR_real_network_8x7.png}
    \label{pr_real}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{REC}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_REC_real_network_8x7.png}
    \label{rec_real}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $50/50$ opinion distribution. The Last.fm on-line social network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $50/50$ opinion distribution. The Last.fm on-line social network is compared to corresponding network models: (i) WS with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) SBM-WS with 53 communities of size 151, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$; (iii) SBM with 53 communities of size 151 with $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$. \textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Each data point is the average of $50$ independent simulations. Bars represent the standard deviation $\sigma$.}
\label{echo_vs_fracRes_real}
\end{figure}
\noindent
The REC filtering algorithm enforces an increase in the formation of echo chambers when a small fraction of stubborn actors is introduced. If this fraction is further increased, the formation of echo chambers starts to decrease. The increase is the strongest for the high clustering network models (WS, SBM-WS). Even though the Last.fm network has a similar clustering as these high clustering models, the formation of echo chambers and the corresponding increase in that formation are significantly smaller. As for the PR filtering algorithm, this is probably a consequence of the heterogeneous degree pattern of the Last.fm network. The results for the Last.fm network coincide with those of the SBM. This seems to suggest that the hampering in the formation of echo chambers due to a heterogeneous degree distribution is of a similar size as the hampering due to the lack of clustering.

@@

Results for the PGP trust network are compared to results for network models with similar average degree, clustering and/or modularity. The used network models are, all with $N = 10680$: (i) the Watts-Strogatz model with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) the stochastic block model with 60 communities of size 178, $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$; (iii) the stochastic Watts-Strogatz block model with 60 communities of size 178, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$.\\
\newline
In Fig. \ref{echo_vs_fracRes_real_PGP} the formation of echo chambers versus the fraction of completely stubborn actors (fracRes) is studied for the real-world PGP trust network and for the corresponding network models. This is done for both the PR and REC filtering algorithms. \\
\newline
The PR filtering algorithm (Fig. \ref{pr_real_PGP}) shows a decrease in the formation of echo chambers with an increasing fraction of completely stubborn actors for both the PGP trust network and the corresponding network models. The real-world PGP network has a significantly smaller formation of echo chambers compared to the network model with similar clustering and modularity (SBM-WS). This is most likely due to the heterogeneous degree pattern of the PGP network. \\
The REC filtering algorithm (Fig. \ref{rec_real_PGP}) shows an increase in the formation of echo chambers with an increasing fraction of stubborn actors, followed by a decrease if the fraction of stubborn actors is further increased. This is visible for both the PGP trust network and the network models. Again, the PGP network has a smaller formation of echo chambers than the network model with similar clustering and modularity (SBM-WS), due to its heterogeneous degree distribution. The data points for the SBM and the PGP trust network coincide, which may suggest that the hampering in the formation of echo chambers due to a heterogeneous degree distribution is of a similar size as the hampering due to a lack of clustering.
% compare powerlaw community distribution??

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{PR}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_PR_real_network_PGP_8x7.png}
    \label{pr_real_PGP}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{REC}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_REC_real_network_PGP_8x7.png}
    \label{rec_real_PGP}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $50/50$ opinion distribution. The PGP trust network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $50/50$ opinion distribution. The PGP trust network is compared to corresponding network models: (i) WS with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) SBM-WS with 60 communities of size 178, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$; (iii) SBM with 60 communities of size 178 with $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$. \textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Each data point is the average of $50$ independent simulations. Bars represent the standard deviation $\sigma$.}
\label{echo_vs_fracRes_real_PGP}
\end{figure}

@@

It must be emphasized that there are limitations to our model and methods. First, the simpleness of the model, which on one hand can be seen as a strength, may also be one of the weaknesses. Real-life social behavior and opinion dynamics on social media may be far from simple. As a consequence, there may be many dynamics of real-life opinion formation that are not well-captured by our model. More empirical data is therefore needed to compare the results, obtained from our model, against. Other limitations may be the rather small system sizes that were used in our simulations ($N = 10^3$). It is unclear how the system size affects the results and it is recommended that the model is tested on both larger network models and real-world networks. Another possible shortcoming is the binary nature of the model. There are real-life cases in which people have to choose between two competing opinions. These situations, however, do not cover the complete range of possible opinion dynamics situations. It would be interesting to run the model on situations where more than two opinions are allowed or to switch to continuous opinion dynamics. We also did not differentiate between the agents in our model, except for non-stubborns versus stubborns. This combined with the homogeneous degree distribution of our networks models may be another limitation. Real-life social media platforms display a heterogeneous contact pattern, where the `hubs' are often more influential than other users, not only by their larger reach, but also by their status. This may impact the spread of opinions \cite{Fan2020}\cite{Wang2019}. We could adapt our model to mimic this real social media behavior more closely. Therefore, a heterogeneous network model should be introduced, combined with an influence parameter. This parameter should be used to determine the influence of an agent on others: the opinion of a more influential agent is easier adopted by the others. In the model, the high degree nodes should then be made more influential. As a final limitation, we mention the fixed stubbornness of the agents. Even though the stubbornness parameter $r$ could be individualized to each node, we did not include the possibility to be more stubborn towards opinions coming from one friend than from another. This idea may be linked to the influence of certain friends/acquaintances: somebody may adopt another viewpoint more easily from certain friends than from others.\\

@@

Simple toy models were implemented to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions. Our goal was to get a better understanding of the impact of social media and algorithmic personalization on phenomena such as polarization and formation of echo chambers (opinion clusters) in society. \\
\newline
We found that topological correlations such as high clustering strongly enhance the formation of opinion clusters. Beside clustering, community structure also favors the formation of echo chambers. Real-world networks have a lower formation of echo chambers than network models with similar clustering and modularity. The heterogeneous contact pattern of these real-world networks is most likely the cause of this hampering. If people are not stubborn, we found that semantic filtering and hence, a homogeneous exposure to information increases polarization in the system. This is an intuitive result: if you shield people from opposing views and strengthen them in their own beliefs, it feels natural that the system will evolve towards a more polarized state. It is then expected that a more heterogeneous exposure to information reduces this polarization. We found that this is indeed the case if people are not stubborn. If, on the contrary, people have some level of stubbornness, we found the surprising result that homogeneous exposure to information may be more protective against opinion clusters than heterogeneous exposure. This is an important result, since it may change our beliefs/assumptions that algorithmic personalization is at the root of the increase in polarization and extremism often seen on social media. More research on this topic is definitely needed in the near future. Finally, we found that distributing the opinions in communities instead of randomly distributing them may hinder the convergence towards the majority opinion in case of semantic filtering, if the modularity is high enough.\\
\newline
Our results strongly suggest that the effects of algorithmic personalization do not stand on their own. Instead, the interplay with certain network topologies may enhance or hamper these effects. Beside that, individual resistance strongly affects the impact of these filtering algorithms. It would therefore be incorrect to only point to algorithmic personalization as the cause of polarization. It is the interplay between many different aspects of the system (such as topological correlations, distribution of the opinions, individual resistance, content curration) that may potentially enhance or hamper the formation of opinion clusters.
% also discuss need to investigate effect of changing N, S and p_act
\newpage
